---
title: Enhance quantitative analysis of EPMA maps with QntMap
output: github_document
---

[![Travis build status](https://travis-ci.org/atusy/qntmap.svg?branch=master)](https://travis-ci.org/atusy/qntmap)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/atusy/qntmap?branch=master&svg=true)](https://ci.appveyor.com/project/atusy/qntmap)
[![Coverage status](https://codecov.io/gh/atusy/qntmap/branch/master/graph/badge.svg)](https://codecov.io/github/atusy/qntmap?branch=master)
[![license](https://img.shields.io/badge/license-MIT-blue.svg)]()


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(knitr)
```

# Overview

This package generates mass concentration maps and phase distribution maps based on X-ray mapping data and spot analysis data from EPMA.

See "[How to]" for a usage and [Yasumoto et al. (2018)](https://doi.org/10.2138/am-2018-6323CCBY) for implementations.

Current version supports data from JEOL-style EPMA.

# Installation

Copy & paste a following command to R.

```{r install}
source("https://install-github.me/atusy/qntmap")
```

# How to

1. [EPMA analysis] (spot before map)
1. [Export data] from EPMA to PC
1. [Run QntMap on R] for data processing.

Details below.

## EPMA analysis

Conversion is performed by utilizing spot analysis data as internal standards.
Thus, [spot analysis][Spot analysis] must be done prior to [mapping][Mapping].

### Spot analysis

- Analytical conditions
    - Same as those conventionally applied in your lab.
    - Use wavelength-dispersive X-ray spectrometer
- Spots to be analyzed
    - **20 spots per phase** in the area to be mapped (the more is better).
    - It is better but not necessary to quantify grains larger than mapping probe diameter.
    - Make sure at least **20 spots per element** are analyzing grains larger than mapping probe diameter.
- Identify phases in comment
    - Give same comments on the same phase with the similar compositions.
      - e.g., quartz, plagioclase, garnet-core, garnet-rim, ...
    - An alternative is to use external file later.

### Mapping

- Analytical conditions
    - Acceralating voltage must be same as that in spot analysis.
    - Probe diameter should be larger than that in spot analysis. 
    - Probe current is recommended to be 100 nA following Lanari et al. (2014).
    - Dwell time is recommended to be 0.1 - 0.3 sec following Lanari et al. (2014).

### Example of analytical conditions

|                     | Spot  | Map     | Comment                         |
|:--------------------|------:|--------:|:--------------------------------|
|Acceralating Voltate | 15 kV | 15 kV   | Must be same in spot and map    |
|Probe diameter       | 3 μm  | 20 μm   | Must be smaller in spot than map|
|Probe current        | 10 nA | 100 nA  |                                 |
|Peak dwell           | 10 sec| 120 msec|                                 |
|Background dwell     |  5 sec| NA      | No need to analyze in map       |


## Export data {#Export}

1. **ASCII conert** mapping data into matrix format,
   and save the result in the directory
   where raw data is stored (e.g., `.map/1`).
1. Export whole directory of analysis containing 
   `.map` directory and `.qnt` directory

### File structure

A minimal example.
See links for descrptions.
`*` is a wild card.

- .map/\*/
    - \*_map.txt
    - \*.cnd
- .qnt/
    - [.cnd/elemw.cnd](#elemwcnd)
    - [bgm.qnt](#bgmqnt-bgpqnt-pkintqnt-netqnt)
    - [bgp.qnt](#bgmqnt-bgpqnt-pkintqnt-netqnt)
    - [elem.qnt](#elemqnt-elintqnt)
    - [elint.qnt](#elemqnt-elintqnt)
    - [mes.qnt](#mesqnt)
    - [net.qnt](#bgmqnt-bgpqnt-pkintqnt-netqnt)
    - [peak.qnt](#peakqnt)
    - [stg.qnt](#stgqnt)
    - [wt.qnt](#wtqnt)

#### elemw.cnd

includes dwell time for peak and background, 
relative positions of backgrounds.

In some case, this file is incomplete or missing, and needs to be prepared manually (e.g., https://gist.github.com/atusy/f1577b67b8874c9e915941c0725d0e22).

#### bgm.qnt, bgp.qnt, pkint.qnt, net.qnt

include
background (minus and plus), peak, and net intensities of each analysis.
In some environemts, `pkint.qnt` is missing, but is complemented within QntMap.

#### elem.qnt and elint.qnt

include name of elements 
data correction ("elem.qnt"; oxide or metal in ZAF),
counting intensities ("elint.qnt").

#### mes.qnt

includes probe current of each analysis.

#### peak.qnt

includes positions of peak intensities of each element of each analysis.

#### stg.qnt

includes coordinates and comments of each analysis.

#### wt.qnt

includes mass concentrations of each analysis.

## Run QntMap on R

For data processing

### Interactive mode

Follow instructions shown by running the following code.

```{r interactive}
library(qntmap)
qntmap()
```

As a result, phase identification result is saved in "`clustering`" directory and mass concentration data as csv files in "`qntmap`" directory both under the directory contaning mapping data.

Note that interactive mode has limited functions.
Use [manual mode][Manual mode] for full functionality.

### Manual mode

A work-flow with example dataset is available at
https://atusy.github.io/qntmap/articles/basic.html .

```{r manual}
library(qntmap)

# Required parameters
wd <- '.' # path to the working directory
dir_map <- '.map/1' # relative/absolute path to the directory containing ascii converted X-ray map files (1_map.txt, 2_map.txt, and so on)"
dir_qnt <- '.qnt' # relative/absolute path to the directory containing .qnt files (pkint.qnt, net.qnt, and so on)"


# Optional parameters

## A character vector to specify phases tend to be smaller than mapping probe diameter
fine_phase <- NULL 

## A csv file indicating name of the phase of n-th quantitative point analysis.
## The file path is absolute or relative to `dir_qnt`.
## If NULL, names are assumed to be specified in comments during EPMA analysis.
phase_list <- NULL 

# Run analysis

# Set working directory
setwd(wd)

# Load mapping data
# Change value of DT (dead time in nanoseconds) depending on EPMA.
# 1100 ns is a value applied by JEOL JXA-8105.
xmap <- read_xmap(wd = dir_map, DT = 1100)

# Compile quantitative data
qnt <- read_qnt(wd = dir_qnt, phase_list = phase_list, renew = TRUE)
## Check 'phase_list0.csv' under 'dir_qnt' to see if name of phases are provided properly.
## If not, modify the csv file and specify the path of modified one to `phase_list` in "Optional parameters" section and rerun the above code.

# Determine initial cluster centers
centers <- find_centers(xmap = xmap, qnt = qnt, fine_phase = fine_phase)
## Check 'centers0.csv' under the `wd` and modify on demand.
## If modified, assign content of the modified csv file by running
## centers <- data.table::fread('path to the modified csv file')

# Phase identification
# Assign group_cluster = TRUE if you want to integrate same phases subgrouped by suffix after '_' 
# (e.g., garnet_a and garnet_b are integrated to garnet if TRUE)
cls <- cluster_xmap(xmap = xmap, centers = centers, group_cluster = FALSE)

# Quantify X-ray maps
qmap <- quantify(
  xmap = xmap, qnt = qnt, cluster = cls, fine_phase = fine_phase
)
## Resulting files are saved in `qntmap` directory` under `dir_map`.

# Summarize result
summary(qmap)
## This shows minimum, lower quantile, median, mean, upper quantile, and maximum values of variables.
```
