---
title: "Add phases after EPMA analysis"
author: "YASUMOTO Atsushi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- © 2018 JAMSTEC -->

```{r setup, include = FALSE}
wd <- tempdir()
knitr::opts_chunk$set(
  root.dir = wd,
  fig.width = 7,
  fig.height = 7,
  collapse = TRUE,
  comment = "#>"
)
library(qntmap)
library(dplyr)
library(pipeR)
library(ggplot2)
```

# Intro

This vignette introduce a way to quantify X-ray map even if, in accident, 
there are some phases without spot analysis. 
For accurate analysis, I highly recommend careful observation prior to EPMA analysis, 
so that as less phases as possible being missed.

Please read [basic.html](basic.html) before hand.

# Preparation

This section prepares package, files, and modify sample data.
An original sample dataset is same as that used in ["Basic usage"](basic.html).
However, the dataset will be modified to let a phase be not analyzed by EPMA spot analysis.
This is why "[Specify directories containing data]" and "[Load Data]" belongs to "[Preparation]" despite
they belongs to [Analysis section of "Basic usage"](basic#Analysis).

## Load package

```{r, eval = FALSE}
library(qntmap)
```

## Move to working directory

```{r eval = FALSE}
wd <- tempdir() # Arbitrary
setwd(wd)
```

## Copy example data

or any original data.

```{r}
file.copy(
  from = system.file('extdata', 'minimal', package = 'qntmap'),
  to = wd,
  recursive = TRUE
)
```

`TRUE` indicates files are successfully copied. 
Check if really files are copied by `dir()`.

```{r}
dir(file.path(wd, 'minimal'), recursive = TRUE, all.files = TRUE)
```

## Specify directories containing data

```{r}
dir_map <- file.path(wd, 'minimal/.map/1')
dir_qnt <- file.path(wd, 'minimal/.qnt')
```


## Load data

```{r}
xmap <- read_xmap(dir_map)
qnt <- read_qnt(dir_qnt)
```

## Check quantified data

In the original dataet, 20 spots are analyzed by EPMA on both olivine and quartz.

```{r}
table(qnt$cnd$phase)
```

## Modify data

Let's delete quartz from dataset and reload.

After `read_qnt()` is performed, a file `phase_list0.csv` is created under `dir_qnt`.
This file consists of 3 columns `id`, `phase`, and `use`.

**Do not** edit `id`.

**Correct** `phase` in the analyzed phases are not expected ones, 
or in case some phase needs be sub-grouped due to large compositional variation.
For example, substitute Ol by Ol_Fe and Ol_Mg based on Fe/Mg ratio.

**Correct** `use` in case there are bad results.

This time, let `use` be `FALSE` if `phase` is Qtz, and retry `read_qnt`

To do it on R, execute codes below.
Be sure to specify a path to modified csv file to `phase_list` parameter of `read_qnt`.
Also, `renew` parameter of `read_qnt` must be `TRUE`.

```{r}
phase_list <- read.csv(file.path(dir_qnt, 'phase_list0.csv'))
phase_list$use[phase_list$phase == 'Qtz'] <- FALSE
write.csv(phase_list, file.path(dir_qnt, 'phase_list_no_qtz.csv'))
qnt <- read_qnt(dir_qnt, phase_list = file.path(dir_qnt, 'phase_list_no_qtz.csv'), renew = TRUE)
```

After editing, only olivine is said to be quantified.

```{r}
table(qnt$cnd$phase)
```

```{r, message = FALSE, echo = FALSE}
epma <- tidy_epma(qnt, xmap) %>>%
  filter(elm == elm[[1]]) %>>%
  select(x_px, y_px, phase)
plot(xmap, 'Si', interactive = FALSE) +
  scale_fill_gradient(low = 'black', high = 'white') +
  geom_point(
    aes(y_px, x_px, colour = phase),
    data = epma, inherit.aes = FALSE
  )
```

# Analysis

## Cluster analysis

### Initialize cluster centers

```{r}
centers <- find_centers(xmap, qnt)
centers
```

### Compare the initial centroids with X-ray map plotted with heatmap

```{r}
plot(xmap, 'Si', interactive = TRUE)
```

In this case, an analysist will notice there is something other than olivine.
Let's mouse over the **interactive map** and one will see one of the coordinates of quartz (e.g., x = 18, y  = 28).
Keep it in your note.

### Add initial centers

#### on R

Add initial centers by using `add_centers`.

```{r}
centers <- add_centers(centers = centers, xmap = xmap, x = 18, y = 28, p = 'Qtz')
```

Then, Qtz is properly added to `centers`.

```{r}
print(centers)
```
#### on Spreadsheet program + R

- Open `centers_initial0.csv` in the current directory (try `getwd()` on R if unknown).
- Edit it
- Save it
- Reload on R by `centers <- read.csv('path to an updated csv file')`

### Cluster and quantify

```{r fig.show = 'hide'}
cluster <- cluster_xmap(xmap, centers)
qmap <- quantify(xmap, qnt, cluster)
```

```{r}
summary(cluster)
summary(qmap)
```

```{r include = FALSE}
unlink(c('centers0.csv', 'center_add.csv'))
```
