---
title: "Basic usage"
author: "Atsushi YASUMOTO"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Basic usage}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<!-- © 2018 JAMSTEC -->

```{r setup, include = FALSE}
wd <- tempdir()
knitr::opts_chunk$set(
  root.dir = wd,
  fig.width = 7,
  fig.height = 7,
  collapse = TRUE,
  comment = "#>"
)
library(qntmap)
library(dplyr)
library(pipeR)
library(ggplot2)
library(stringr)
library(purrr)

formals(read_xmap)$renew <-
  formals(read_qnt)$renew <- 
  TRUE
```

# Intro

This vignette introduce a basic procedure of analysis using `qntmap` package using example data.

See https://atusy.github.io/qntmap/#epma-analysis for preparing EPMA data.

See https://doi.org/10.2138/am-2018-6323CCBY for implementations.

# Preparation

## Load package

```{r load-qntmap, eval = FALSE}
library(qntmap)
```

## Move to working directory

```{r setwd, eval = FALSE}
wd <- tempdir() # Arbitrary
setwd(wd)
```

## Copy example data

or any original data.

```{r copy-files}
file.copy(
  from = system.file('extdata', 'minimal', package = 'qntmap'),
  to = wd,
  recursive = TRUE
)
```

`TRUE` indicates files are successfully copied. 
Check if really files are copied by `dir()`.

```{r file-list}
dir(file.path(wd, 'minimal'), recursive = TRUE, all.files = TRUE)
```

# Analysis

## Specify directories containing data

```{r specify-directories, include = FALSE, eval = TRUE}
dir_map <- file.path(wd, 'minimal/.map/1')
dir_qnt <- file.path(wd, 'minimal/.qnt')
```

``` r
dir_map <- 'minimal/.map/1'
dir_qnt <- 'minimal/.qnt'
```


## Load data

```{r, message = FALSE}
xmap <- read_xmap(dir_map)
qnt <- read_qnt(dir_qnt, renew = TRUE)
```

This is an artifact example data which contains olivine and quartz within the mapping area.
The phases are quantified 20 points each.

```{r, message = FALSE, echo = FALSE, warning = FALSE}
epma <- tidy_epma(qnt, xmap) %>>%
  filter(elm == elm[[1]]) %>>%
  select(x_px, y_px, phase)
plot(xmap, 'Si', interactive = FALSE) +
  scale_fill_gradient(low = 'black', high = 'white') +
  geom_point(
    aes(y_px, x_px, colour = phase), data = epma, inherit.aes = FALSE
  ) +
  guides(
    fill = guide_colourbar(barheight = grid::unit(1, "npc") - unit(10, "line"))
  )
```


## Plot X-ray map data as heatmap

Plots are by default **interactive** with webui.
Selection of elements can be done on webui.

```{.r}
plot(xmap)
# A following example is non-interactive mode run by
# plot(xmap, 'Mg', interactive = FALSE)
```

```{r, echo = FALSE}
plot(xmap, 'Mg', interactive = FALSE)
```


## Cluster analysis

### Initialize cluster centers

Although this step is just a guessing initialization, 
compare the output values with plots above 
to see if the centers are well guessed.

```{r}
centers <- find_centers(xmap, qnt)
centers
```

### Run cluster analysis {#run-cls}

The result is saved as a binary file and as png format images in `clustering` directory below the directory containing mapping data.

For a quick look, `plot()`{.r} function is also available.

```{r, eval = TRUE, fig.keep = "last"}
cluster <- cluster_xmap(xmap, centers)
plot(cluster, interactive = FALSE)
```


### Summarize cluster analysis

`summary` function gives abundance ratios of phases in the map.

```{r}
summary(cluster)
```

## Quantify

```{r}
qmap <- quantify(xmap = xmap, qnt = qnt, cluster = cluster)
```

### Plot quantified map

Just like [X-ray map](#Plot X-ray map).

```{r}
plot(qmap, 'SiO2', interactive = FALSE)
```

### Summarize quantified map

```{r}
summary(qmap)
```

Note also that this summary does not correct densities of phases.

```{r, include = FALSE}
V <- unclass(table(cluster$cluster))
V <- V / sum(V)
E <- mean(qmap)
E <- setNames(round(E[['Whole area']], 2), E[['Element']])
```

According to compositions of olivine (57.29% MgO and 42.71 wt% SiO~2~),
and of quartz (100% SiO~2~),

$$
E(MgO_{bulk}) = 
  57.29 \times `r V['Ol']` + 0 \times `r V['Qtz']` = 
  `r 57.29 * V['Ol']` \simeq 
  `r E['MgO']` \mathrm{,}
$$

and

$$
E(SiO2_{bulk}) = 
  42.71 \times `r V['Ol']` + 100 \times `r V['Qtz']` = 
  `r 42.71 * V['Ol'] + 100 * V['Qtz']` \simeq
  `r E['SiO2']` \mathrm{,}
$$

the above summary is adequate.

### Summary based on mask images

For another way to summarize the result of quantified map, `mean` function is defined.

A simple usage gives nothing more than `mean` column of `summary()`.

```{r}
mean(qmap)
```

However, by using mask image, 
`mean` values of each element are calculated for each mask areas with same colors.

Let's use an image from cluster analysis [above](#run-cls).

Input path to the image to `segment` function.

```{r img-path, include = FALSE}
img <- dir(
    file.path(dir_map, "clustering"), pattern = "_map.png", full.names = TRUE
  )[[1]]
i <- segment(img)
```

``` r
i <- segment(dir("`r img`"))
```


Then, input the result of `segment()` to `index` parameter of `mean()`.

```{r}
mean(qmap, index = i)
```

As black (`#000000`) corresponds to olivine cluster,
and white (`#FFFFFF`) corresponds to quartz cluster,
the average compositions of each clusters are nearly equal 
to ideal compositions of olivine and quartz.

This feature is useful to find average compositions of each minearls, 
local bulk compositions of domains (layer, symplectite, and so on).

Note that mask image must be png image 
whose **width and height are same as mapping data**.

```{r include = FALSE}
unlink(c('centers0.csv', 'phase_list0.csv'))
```

