---
title: "Inherit parameters from previously quantified maps"
author: "YASUMOTO Atsushi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inherit parameters from previously quantified maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(dplyr)
library(knitr)
library(qntmap)
```

<!-- © 2018 JAMSTEC -->

# Introduction

This article introduces a way to quantify X-ray maps without quantitatively analyzed spots by 
inheriting parameters from a previously quantified X-ray map (See figure below).

```{r sample-with-two-mapping-areas, echo = FALSE, fig.width = 6, fig.height = 3}

set.seed(1)
spots <- runif(40) %>%
  matrix(ncol = 2, dimnames = list(NULL, c("x", "y"))) %>%
  as.data.frame
rect <- data.frame(xmin = c(0, 3), xmax = c(1, 4), ymin = 0, ymax = 1)

ggplot(spots) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), data = rect) +
  geom_point(aes(x, y)) +
  annotate("text", x = 2, y = -Inf, label = "Sample", vjust = -1, size = 8) +
  annotate("text", x = c(0.5, 3.5), y = 1, vjust = -.3, size = 5, label = c(
    "(a) Mapped area\nwith spots",
    "(b) Mapped area\nwithout spots"
  )) +
  xlim(-1, 5) +
  ylim(-1.5, 2.5) +
  theme(
    axis.ticks=element_blank(),
    axis.ticks.length=unit(0, "null"),
    axis.text=element_blank(),
    axis.title=element_blank(),
    axis.line=element_blank(),
    panel.background=element_rect(fill = "gray",colour = NA),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.spacing=unit(c(0,0,0,0), "null"),
    plot.margin=unit(c(0,0,0,0), "null"),
    legend.position="none",
    legend.margin=margin(0,0,0,0, "null"),
    NULL
  ) +
  NULL
  
```


# Requirements

- `qntmap 0.3.2` or later
- Referencial quantified maps from a single area.
    - The maps must be constructed by `quantify(saving = TRUE)`
      under `qntmap 0.3.2` or later.
    - To construct a referencial maps, read ["Basic usage"](basic.html).
- X-ray maps to be quantified.
    - The new maps must be analyzed under the same conditions to the reference maps.
    - The new maps must comprise mineral assemblage which is a subset of the reference maps.

# Instructions

## Load qntmap package

```{.r}
library(qntmap)
```

## Read X-ray mapping data to be quantified

```{.r}
xmap <- read_xmap(".map/2 for example")
```

## Cluster analysis

### Initialize cluster centers

Suppose available quantified maps are constructed by following 
["Basic usage"](basic.html), `centers_initial0.csv` is available as well.
Read the csv file by following codes.

```{.r}
centers <- read.csv("example/centers_initial0.csv")
centers
```

```{r find-centers, echo = FALSE}
dirs <- file.path(system.file('extdata', 'minimal', package = 'qntmap'), c(".map/1", ".qnt"))
xmap <- read_xmap(dirs[[1]], saving = FALSE)
qnt <- read_qnt(dirs[[2]], saving = FALSE)
kable(find_centers(xmap, qnt, saveas = FALSE))
```

The above is a example format of initial centers given by 
sample data introduced in ["Basic usage"](basic.html).
Edit the csv file manually if initial centers require modifications, or
if there are any phase abscent in new maps to be quantified.

### Run cluster analysis

```{.r}
cluster <- cluster_xmap(xmap, centers)
```

## Quantification

When available quantified mapes are compsoed by `quantify()` function,
`parameters.csv` is saved in the `qntmap` directory within the directory
storing X-ray mapping data.
Specify a file path of `parameters.csv` as `fix` parameter of 
`quantify()` function.

```{.r}
quantify(xmap = xmap, qnt = NULL, cluster = cluster, fix = "example/qntmap/parameters.csv")
```

Note that `qnt` paramter is ignored when `fix` parameter is specified.
